\bf Concurrent Data Assimilation for Model-guided Learning of Cardiac Potassium Channel Activities

### 2023-07-10 writing/draft_new.tex
2nd revision

\begin{abstract}
Potassium channels (K\textsubscript{v}) are responsible for repolarizing the action potential in cardiomyocytes. There is a variety of K\textsubscript{v} isoforms and corresponding currents (e.g., I\textsubscript{Kto}, I\textsubscript{Kslow1}, I\textsubscript{Kslow2}) that contribute to different phases of repolarization. Because only the sum of their activities can be measured in the form of currents (I\textsubscript{Ksum}), there is a need to decompose and estimate K\textsubscript{v} currents and kinetics. Most existing studies separate and make inference of K\textsubscript{v} kinetics via curve-fitting procedures (e.g., tri-exponential fitting), but are limited because: 1) curve-fitting decomposition only relies on the shape of K\textsuperscript{+} current traces, which does not discern the underlying kinetics and interactions of K\textsubscript{v} isoforms; 2) I\textsubscript{Ksum} traces can only be fitted for one clamp voltage at each time, and estimated information is analyzed in a population-averaged way later. Here, we develop a novel concurrent data assimilation method that calibrates biophysics-based subject-specific computer models to decompose and delineate kinetics of K\textsubscript{v} isoforms with multiple voltage-clamp responses simultaneously. The proposed method is evaluated and validated with whole-cell I\textsubscript{Ksum} recordings from wild-type and chronically glycosylation-deficient cardiomyocytes. Experimental results show that the proposed method effectively handles multiple responses in the voltage-clamp protocol and describes the glycosylation-conferred perturbations observed experimentally to various K\textsubscript{v} isoforms. In addition, we develop a graphical-user-interface (GUI) application that provides an enabling tool to biomedical scientists. The proposed method and pertinent software are shown to have strong potential to study K\textsubscript{v} kinetics in various heart diseases.
\end{abstract}

@@

\section{Introduction}
Potassium channels (K\textsubscript{v}) play critical roles in the electrical conduction system of the heart, particularly in the repolarization phase of the action potential (AP). The AP is a change of membrane potential over time, representing the net electrical activity in a cardiomyocyte, and the AP shape and duration are primarily determined by K\textsubscript{v} isoforms. As shown in Figure~\ref{fig:ap_currents}, the different phases of AP repolarization in mouse cardiomyocytes are the result of the coordinated activity of a variety of K\textsubscript{v} isoforms (e.g., K\textsubscript{v}4.2, K\textsubscript{v}1.5, and K\textsubscript{v}2.1) and their corresponding currents (e.g., I\textsubscript{Kto}, I\textsubscript{Kslow1}, and I\textsubscript{Kslow2}). There are other major voltage-gated ion channels (VGICs), such as for Na\textsuperscript{+} (Na\textsubscript{v}) and Ca\textsuperscript{2+} (Ca\textsubscript{v}), that also contribute to the AP. The Na\textsubscript{v} are primarily responsible for the AP upstroke in the depolarization phase, while the Ca\textsubscript{v} control the cellular contraction. Aberrant activities of ion channels can significantly impact the AP and lead to fatal arrhythmias. Using whole-cell voltage-clamp recording methods, only the collective activities of K\textsubscript{v} isoforms can be measured reliably as the sum of all K\textsuperscript{+} currents (I\textsubscript{Ksum}), and these K\textsubscript{v} isoforms have overlapping biophysical properties (i.e., voltage dependence of gating and kinetics), which complicate the ability to separate one type of I\textsubscript{K} from another \citep{brouillette2004functional}. K\textsubscript{v} activities can be altered in various cardiomyopathies \citep{tristani2001molecular,giudicessi2012potassium}, thus there is an urgent need to decompose I\textsubscript{Ksum} more rigorously into individual K\textsuperscript{+} currents and estimate gating kinetics of K\textsubscript{v} isoforms to understand their pathological roles by comparing their activities in healthy versus diseased hearts and cardiomyocytes.
\begin{figure}[!ht]
    \centering
    \includegraphics{figs/ap_currents.pdf}
    \caption{(a) Ventricular action potential in mouse cardiomyocytes and (b) diverse K\textsubscript{v} isoforms and their currents contributing to the AP shape.}
    \label{fig:ap_currents}
\end{figure}

@@

This paper presents a new approach of subject-specific concurrent data assimilation for model-guided learning of K\textsubscript{v} isoforms. First, computer models of K\textsubscript{v} and their currents are designed with parameters that control the kinetic rates and, in turn, determine the currents. It is worth mentioning that K\textsuperscript{+} currents are generated based on the mathematically simulated gating mechanism as opposed to the curve-fitting procedure, which only mimics the shape of the currents. Second, we perform a sensitivity analysis of the kinetic parameters using fractional factorial designs to identify the parameters that have significant effects on the current generation. Several markers are defined that best represent the characteristics of the K\textsuperscript{+} currents in different perspectives. Third, a calibration routine is proposed to adjust parameters selected by the sensitivity analysis to couple \textit{in-silico} models with \textit{in-vitro} data of I\textsubscript{Ksum} recordings from the same cell concurrently. Last, low-dimensional embedding is utilized to visualize calibrated parameters in a collective way across the cells. The proposed methodology is evaluated and validated by comparing I\textsubscript{Ksum} recordings on healthy control versus chronically glycosylation-deficient cardiomyocyte. Glycosylation is a co/posttranslational modification that is critical for protein functions including activities of K\textsubscript{v} and other VGICs \citep{ohtsubo2006glycosylation,ednie2012modulation}. Our \textit{in-vitro} experiments showed that chronic reduction in cardiomyocyte N-glycosylation modulates VGIC activities and contribute to both electrical and contractile dysfunction \citep{ednie2013sialicNav1,ednie2015sialicKv,ednie2019reduced,ednie2019reduced2}. In fact, these studies also showed that reduced cardiomyocyte complex N-glycosylation is sufficient to cause dilated cardiomyopathy (DCM), which is the third most common cause of heart failure and the major reason for heart transplantation \citep{weintraub2017dilated}.

Experimental results show that the proposed method provides reliable estimations of K\textsubscript{v} currents with underlying kinetics and successfully captures cellular variability. For example, it verifies the experimental data that showed the significant reduction in the current magnitude and elongated decaying for the glycosylation-deficient cardiomyocyte group. Additionally, computational modeling estimates channel kinetics, which was not possible in the electrophysiological experiments alone. Cellular variability is visualized by low-dimensional embedding of calibration parameters into 3D space, which shows one of the powers of systematic computer models that allow one to predict how changes in VGICs impact cellular functions. Further, we develop a graphical-user-interface (GUI) application to make the proposed method accessible to biomedical scientists for investigating K\textsubscript{v}-related channelopathy that is available as online supplemental material. The proposed method shows strong potential for modeling the kinetics and gating mechanism of K\textsubscript{v} to study heart diseases. Our contributions are summarized as follows:
\begin{itemize}
    \item We model the underlying kinetics of K\textsubscript{v} isoforms from I\textsubscript{Ksum} recordings, which enables to simulate individual K\textsuperscript{+} currents based on biophysical principles,
    \item We develop a subject-specific concurrent data assimilation routine that learns I\textsubscript{Ksum} recordings measured from the same cell at multiple membrane potentials simultaneously rather than a single I\textsubscript{Ksum} trace independently to model cellular-level dynamics caused by voltage dependence, 
    \item The prediction uncertainty is quantified and the distribution of kinetic parameters is visualized using low-dimensional embedding,
    \item A GUI application is provided to make the proposed method an enabling tool for biomedical scientists.
\end{itemize}

\section{Research Background}
\subsection{Data Assimilation and Calibration of Cardiac Models}
The functionality of the heart to pump blood through the body is controlled by electrical signals generated by the organ itself. Cardiac electrophysiology has contributed significantly to our understanding of the heart and disease-related modifications, such as congenital disorders of glycosylation (CDG). Advancements in laboratory techniques allow researchers to study molecular-level activities in cardiomyocytes via measuring ionic currents conducted through VGICs. For example, whole-cell current recordings are used to show the functions of a certain gene that encodes $\alpha$ subunits of Na\textsubscript{v} in human cardiomyocytes, leading to a small but inherent and chronic risk of acquired arrhythmia \citep{splawski2002variant}; pathophysiological roles of reduced sialylation impacting Na\textsubscript{v} and K\textsubscript{v} activities in mouse cardiomyocytes \citep{ednie2015sialicNav2, ednie2015sialicKv}; modulation of O-glycosylation causing aberrant K\textsubscript{v} activities \citep{schwetz2011sialic}. However, \textit{in-vitro} experiments alone are limited to investigate detailed channel activities. There is a need to assimilate experimental data to estimate information that are not able to be observed directly. Traditionally, a curve-fitting method is used to make inferences, which assumes a functional form of information to estimate (e.g., individual currents of K\textsubscript{v} isoforms) \citep{liu2011dissection}.

However, curve-fitting approaches have limitations in that they only consider the shape of informational functions to infer rather than incorporating underlying kinetics and interactions of various gating processes. Further, as curve fitting works for a single curve independently, the traditional approach is less concerned about the dynamics of a set of multiple responses measured from the same cell (e.g., the coincidental activity of multiple K\textsubscript{v} isoforms). In contrast, computer models consist of mathematical equations describing biophysical properties and physiological functions compatible with experimental observations. Mathematical and computational modeling allows for studying the heart in a quantitative and predictive way \citep{whittaker2020calibration}. Computer models, coupled with experimental observations, provide integrative insights into the data by calibrating model parameters \citep{winslow2011integrative}. For example, models can be used to interpolate processes not directly observed in experiments and extrapolate to novel conditions such as disease-related perturbations \citep{rodriguez2010systems}. From an engineering perspective, computer models of ion channels are dynamic system simulations that represent continuously changing gating kinetics. The dynamic nature of the models complicates the calibration process.

Complex nonlinear structures of cardiac computer models also make calibration difficult. We developed statistical metamodeling and sequential design \citep{du2015statistical}, nonlinear optimization algorithms \citep{du2013silico,du2017silico}, and a heuristic optimization method \citep{kim2022simulation} in our previous \textit{in-silico} studies to cope with the complexity of cardiac models. However, all these studies are based on current-shape parameters via curve fitting, or characteristic curves/statistics from additional curve fitting applied to estimated currents. In addition, only population-average models are derived and calibrated via learning the entire dataset for each healthy and diseased group. Little work has been done to leverage computer models to calibrate and delineate kinetic dynamics of K\textsubscript{v} isoforms with multiple voltage-clamp responses simultaneously.

\subsection{Congenital Disorders of Glycosylation}
Protein glycosylation is one of the most abundant and diverse forms of co/posttranslational modifications that impact essential protein functions, such as modulation of receptor or ion channel activities \citep{ohtsubo2006glycosylation,ednie2012modulation}. A growing number of studies have shown the association between altered glycosylation and heart diseases, such as dilated cardiomyopathy (DCM) and hypertrophic cardiomyopathy \citep{ednie2019reduced2,ohtsubo2006glycosylation}. It is reported that up to 20\% of patients with congenital disorders of glycosylation (CDG), who commonly show modest reductions in protein glycosylation, present with cardiac deficits, including idiopathic DCM \citep{marques2017cardiac}. However, uncovering the underlying pathological mechanisms still remains elusive. We have investigated how regulated glycosylation contributes to heart failure in the context of electrophysiology. Electrical signaling is orchestrated activities of a variety of ion channels and transporters. VGICs are heavily glycosylated, with ~30\% of the channel mass consisting of N-/O-linked glycans \citep{ednie2012modulation}. Glycosylation is a multi-step process and usually ends with sialic acid added. We reported that a saturating, electrostatic effect of negatively charged sialic attached to the terminal of N-/O-glycan branches significantly altered electrical signaling in Na\textsubscript{v} \citep{ednie2013sialicNav1,ednie2015sialicNav2} as well as K\textsubscript{v} \citep{ednie2015sialicKv}. Computational modeling has been used to further investigate the functional role of reduced sialylation in Na\textsubscript{v} and K\textsubscript{v} \citep{du2015statistical, du2017silico}.

\section{Research Methodology}
\subsection{Computer Models of Potassium Channel Isoforms}
Because of the relative ease at which their genetics can be altered, mice have been useful models for studying cardiac electric signaling \citep{milani2014small,nerbonne2004studying}. In ventricular and atrial mouse cardiomyocytes, there are three major components of I\textsubscript{Ksum}: rapidly inactivating transient outward K\textsuperscript{+} currents (I\textsubscript{Kto,f} and/or I\textsubscript{Kto,s}), which are conducted through K\textsubscript{v}4.2 and K\textsubscript{v}1.4 respectively, slowly inactivating delayed rectifier K\textsuperscript{+} currents (I\textsubscript{Kslow1} and I\textsubscript{Kslow2}), which are conducted through K\textsubscript{v}1.5/K\textsubscript{v}2.1, and the non-inactivating steady-state K\textsuperscript{+} current (I\textsubscript{Kss}) \citep{xu1999four}. I\textsubscript{Kto,s} (K\textsubscript{v}1.4) is primarily found in septal cardiomyocytes \citep{bondarenko2004computer}. In this study, we only include I\textsubscript{Kto,f} in I\textsubscript{Kto}, because our \textit{in-vitro} studies focused on the ventricular apex myocytes. However, I\textsubscript{Kto} can be easily modified according to the region of ventricular cardiomyocytes. We model these four dominant K\textsuperscript{+} currents: I\textsubscript{Kto}, I\textsubscript{Kslow1}, I\textsubscript{Kslow2}, and I\textsubscript{Kss}. Figure~\ref{fig:kcurrent_example}(b) illustrates the shape of the primary K\textsuperscript{+} currents and their contribution to the whole-cell I\textsubscript{K} trace (I\textsubscript{Ksum}) given the protocol in Figure~\ref{fig:kcurrent_example}(a) that applies 0 mV voltage pulse from holding potential -70 mV from time $t^{(h)}$ to $t^{(e)}$. Figure~\ref{fig:kcurrent_example}(c) and (d) show a range of voltage steps from -30 mV to 50 mV in 10 mV increments and consequent changes in I\textsubscript{Ksum}. Because of their shape, I\textsubscript{Kss} is assumed as a constant and the other currents an exponential function in the traditional data-driven curve-fitting approach. In this approach, if all four current models are included , I\textsubscript{Ksum} is defined by \eqref{eq:expl_fitting} (i.e., tri-exponential fitting) with the shape parameters such as amplitude $\mathrm{A}_{i}$ and time constant $\tau_{i}$ for $i \in \{\mathrm{Kto}, \mathrm{Kslow1}, \mathrm{Kslow2}, \mathrm{Kss}\}$. In some cases, it is reduced to a bi-exponential function combining I\textsubscript{Kslow1} and I\textsubscript{Kslow2}. 
\begin{figure}[!ht]
    \centering
    \includegraphics{figs/kcurrent_example.pdf}
    \caption{Example of voltage-clamp protocol and K\textsuperscript{+} current traces. (a) Clamp-voltage pulse of 0 mV from the holding potential -70 mV. (b) Dominant K\textsuperscript{+} currents and their contributions to I\textsubscript{Ksum}. (c) Series of clamp-voltage pulses (-30-50 mV in 10 mV increments) from the holding potential -70 mV and (d) consequent I\textsubscript{Ksum} traces.}
    \label{fig:kcurrent_example}
\end{figure}
\begin{equation}
    \label{eq:expl_fitting}
    \mathrm{I}_{\mathrm{Ksum}} = \mathrm{A}_{\mathrm{Kto}}e^{-t/\tau_{\mathrm{Kto}}} + \mathrm{A}_{\mathrm{Kslow1}}e^{-t/\tau_{\mathrm{Kslow1}}} + \mathrm{A}_{\mathrm{Kslow2}}e^{-t/\tau_{\mathrm{Kslow2}}} +  \mathrm{A}_{\mathrm{Kss}}.   
\end{equation}

Unlike the simple current functions, K\textsubscript{v} computer models simulate detailed gating kinetics underlying the currents. We develop mouse K\textsubscript{v} models based on \citep{bondarenko2014compartmentalized,asfaw2020compartmentalized}, using the Hodgkin-Huxley modeling scheme that has been used for various species, such as humans \citep{ten2004model} and rabbits \citep{mahajan2008rabbit}. This type of model consists of two gating variables controlling the channel conductance, and its canonical form is defined by \eqref{eq:hh_ex}
\begin{equation}
    \label{eq:hh_ex}
    \mathrm{I}_{\mathrm{K}} = G_{\mathrm{K}}a^{n}i^{m}(V-E_{\mathrm{K}}),
\end{equation}
where $G_{\mathrm{K}}$ is the maximum conductance, $a^{n}$ and $i^{m}$ are the gating variables for $n,m \in \mathbb{N}$, $V$ is the transmembrane potential, and $E_{\mathrm{K}}$ is the K\textsuperscript{+} Nernst potential. $V-E_{\mathrm{K}}$ implies the driving force of the ion movement. Important components in \eqref{eq:hh_ex} are the gating variables $a$ and $i$, representing the fraction of activation and recovery from inactivation of the channel where $a,i \in [0,1]$. These processes are governed by first-order kinetics and voltage-dependent transition rates $\alpha$ and $\beta$. $\alpha$ is the rate at which a gate in a closed state opens, whereas $\beta$ is the rate at which a gate in an open state closes. Equation~\eqref{eq:gv} shows a schematic relationship of this gating kinetics.
\begin{align}
    \label{eq:gv}
    (1-a)&\xrightleftharpoons[\beta_{a}]{\alpha_{a}}a & (1-i)&\xrightleftharpoons[\beta_{i}]{\alpha_{i}}i.
\end{align}

>>>>

\begin{abstract}
Potassium channels (K\textsubscript{v}) are responsible for repolarizing the action potential in cardiomyocytes. There is a variety of K\textsubscript{v} isoforms and corresponding currents (e.g., I\textsubscript{Kto}, I\textsubscript{Kslow1}, I\textsubscript{Kslow2}) that contribute to different phases of repolarization. Because only the sum of their activities can be measured in the form of currents (I\textsubscript{Ksum}), there is a need to decompose into individual K\textsuperscript{+} currents and their characteristics. Most existing studies separate and make inference of K\textsubscript{v} activities via curve-fitting procedures, but there are limitations such that: 1) curve-fitting decomposition only relies on the shape of K\textsuperscript{+} current traces, which does not discern the underlying kinetics and interactions; 2) I\textsubscript{Ksum} traces can only be fitted for one clamp voltage at each time, and estimated information is analyzed in a population-averaged way later. Here, we develop a novel concurrent data assimilation method that calibrates biophysics-based subject-specific computer models to decompose and delineate kinetics of K\textsubscript{v} isoforms with multiple voltage-clamp responses simultaneously. The proposed method is evaluated and validated with whole-cell I\textsubscript{Ksum} recordings from wild-type and chronically glycosylation-deficient cardiomyocytes. Experimental results show that the proposed method effectively handles multiple responses in the voltage-clamp protocol and describes the glycosylation-conferred perturbations observed experimentally to various K\textsubscript{v} isoforms. In addition, we develop a graphical-user-interface (GUI) application that provides an enabling tool to biomedical scientists. The proposed method and pertinent software are shown to have strong potential to study K\textsubscript{v} kinetics in various heart diseases.
\end{abstract}

@@

\section{Introduction}
Potassium channels (K\textsubscript{v}) play critical roles in the electrical conduction system of the heart, particularly in the repolarization phase of the action potential (AP). The AP is a change of membrane potential over time, representing the net electrical activity in a cardiomyocyte, and the AP shape and duration are primarily determined by K\textsubscript{v} isoforms. The ability of the heart to pump blood through the body in an appropriate rhythm is controlled by electrical signaling. Hence, even modest changes in K\textsubscript{v} activities can significantly affect the AP duration and the QT interval, which lead to fatal heart diseases \citep{ravens2008role}. As shown in Figure~\ref{fig:ap_currents}, the different phases of AP repolarization in mouse cardiomyocytes are the result of the coordinated activity of a variety of K\textsubscript{v} isoforms (e.g., K\textsubscript{v}4.2, K\textsubscript{v}1.5, and K\textsubscript{v}2.1) and their corresponding currents (e.g., I\textsubscript{Kto}, I\textsubscript{Kslow1}, and I\textsubscript{Kslow2}). There are other major voltage-gated ion channels (VGICs), such as for Na\textsuperscript{+} (Na\textsubscript{v}) and Ca\textsuperscript{2+} (Ca\textsubscript{v}), that also contribute to the AP. The Na\textsubscript{v} are primarily responsible for the AP upstroke in the depolarization phase, while the Ca\textsubscript{v} control the cellular contraction. Aberrant activities of ion channels can significantly impact the AP and lead to fatal arrhythmias. Using whole-cell voltage-clamp recording methods, only the collective activities of K\textsubscript{v} isoforms can be measured reliably as the sum of all K\textsuperscript{+} currents (I\textsubscript{Ksum}), and these K\textsubscript{v} isoforms have overlapping biophysical properties (i.e., voltage dependence of gating and kinetics), which complicate the ability to separate one type of I\textsubscript{K} from another \citep{brouillette2004functional}. K\textsubscript{v} activities can be altered in various cardiomyopathies \citep{tristani2001molecular,giudicessi2012potassium}, thus there is an urgent need to decompose I\textsubscript{Ksum} more rigorously into individual K\textsuperscript{+} currents and estimate gating kinetics of K\textsubscript{v} isoforms to understand their pathological roles by comparing their activities in healthy versus diseased hearts and cardiomyocytes.
\begin{figure}[!ht]
    \centering
    \includegraphics{figs/ap_currents.pdf}
    \caption{(a) Ventricular action potential in mouse cardiomyocytes and (b) diverse K\textsubscript{v} isoforms and their currents contributing to the AP shape.}
    \label{fig:ap_currents}
\end{figure}

@@

This paper presents a new approach of subject-specific concurrent data assimilation for model-guided learning of K\textsubscript{v} isoforms. First, computer models of K\textsubscript{v} and their currents are designed with parameters that control the kinetic rates and, in turn, determine the currents. It is worth mentioning that K\textsuperscript{+} currents are generated based on the mathematically simulated gating mechanism. Second, we perform a sensitivity analysis of the kinetic parameters using fractional factorial designs to identify the parameters that have significant effects on the current generation. Several markers are defined that best represent the characteristics of the K\textsuperscript{+} currents in different perspectives. Third, a calibration routine is proposed to adjust parameters selected by the sensitivity analysis to couple \textit{in-silico} models with \textit{in-vitro} data of I\textsubscript{Ksum} recordings from the same cell concurrently. Last, low-dimensional embedding is utilized to visualize calibrated parameters in a collective way across the cells. The proposed methodology is evaluated and validated by comparing I\textsubscript{Ksum} recordings on healthy control versus chronically glycosylation-deficient cardiomyocyte. Glycosylation is a co/posttranslational modification that is critical for protein functions including activities of K\textsubscript{v} and other VGICs \citep{ohtsubo2006glycosylation,ednie2012modulation}. Our in-vitro experiments showed that chronic reduction in cardiomyocyte N-glycosylation modulates VGIC activities and contribute to both electrical and contractile dysfunction \citep{ednie2013sialicNav1,ednie2015sialicKv,ednie2019reduced}. In fact, these studies also showed that reduced cardiomyocyte complex N-glycosylation is sufficient to cause dilated cardiomyopathy (DCM), which is the third most common cause of heart failure and the major reason for heart transplantation \citep{weintraub2017dilated}.

Experimental results show that the proposed method provides reliable estimations of K\textsubscript{v} currents with underlying kinetics and successfully captures cellular variability. For example, it verifies the experimental data that showed the significant reduction in the current magnitude and elongated decaying for the glycosylation-deficient cardiomyocyte group. Additionally, computational modeling estimates channel kinetics, which was not possible in the electrophysiological experiments alone. Cellular variability is visualized by low-dimensional embedding of calibration parameters into 3D space, which shows one of the powers of systematic computer models that allow one to predict how changes in VGICs impact cellular functions. Further, we develop a graphical-user-interface (GUI) application to make the proposed method accessible to biomedical scientists for investigating K\textsubscript{v}-related channelopathy that is available as online supplemental material. The proposed method shows strong potential for modeling the kinetics and gating mechanism of K\textsubscript{v} to study heart diseases. Our contributions are summarized as follows:
\begin{itemize}
    \item We model the underlying kinetics of K\textsubscript{v} isoforms from I\textsubscript{Ksum} recordings, which enables to simulate individual K\textsuperscript{+} currents based on biophysical principles.
    \item We develop a subject-specific concurrent data assimilation routine that learns I\textsubscript{Ksum} recordings measured from the same cell at multiple membrane potentials simultaneously rather than a single I\textsubscript{Ksum} trace independently to model cellular-level dynamics.
    \item The prediction uncertainty is quantified and the distribution of kinetic parameters is visualized using low-dimensional embedding.
    \item A GUI application is provided to make the proposed method an enabling tool for biomedical scientists.
\end{itemize}
The software package of the GUI application, tutorial, and reproducible MATLAB codes for modeling and analysis results are available as supplementary material.

\section{Research Background}
\subsection{Data Assimilation and Calibration of Cardiac Models}
The functionality of the heart to pump blood through the body is controlled by electrical signals generated by the organ itself. Cardiac electrophysiology has contributed significantly to our understanding of the heart and disease-related modifications, such as congenital disorders of glycosylation (CDG). Advancements in laboratory techniques allow researchers to study molecular-level activities in cardiomyocytes via measuring ionic currents conducted through VGICs. For example, whole-cell current recordings are used to show the functions of a certain gene that encodes $\alpha$ subunits of Na\textsubscript{v} in human cardiomyocytes, leading to a small but inherent and chronic risk of acquired arrhythmia \citep{splawski2002variant}; pathophysiological roles of reduced sialylation impacting Na\textsubscript{v} and K\textsubscript{v} activities in mouse cardiomyocytes \citep{ednie2015sialicNav2,ednie2015sialicKv}; modulation of O-glycosylation causing aberrant K\textsubscript{v} activities \citep{schwetz2011sialic}. However, in-vitro experiments alone are limited to investigate detailed channel activities. There is a need to assimilate experimental data to estimate information that are not able to be observed directly. Traditionally, a curve-fitting method is used to make inferences, which assumes a functional form of information to estimate (e.g., individual currents of K\textsubscript{v} isoforms) \citep{liu2011dissection}.

In contrast, computer models consist of mathematical equations describing biophysical properties and physiological functions compatible with experimental observations. Mathematical and computational modeling allows for studying the heart in a quantitative and predictive way \citep{whittaker2020calibration}. Computer models, coupled with experimental observations, provide integrative insights into the data by calibrating model parameters \citep{winslow2011integrative}. For example, models can be used to interpolate processes not directly observed in experiments and extrapolate to novel conditions such as disease-related perturbations \citep{rodriguez2010systems}. From an engineering perspective, computer models of ion channels are dynamic system simulations that represent continuously changing gating kinetics. The dynamic nature of the models complicates the calibration process. Complex nonlinear structures of cardiac computer models also make calibration difficult. We developed statistical metamodeling and sequential design \citep{du2015statistical}, nonlinear optimization algorithms \citep{du2013silico,du2017silico}, and a heuristic optimization method \citep{kim2022simulation} in our previous in-silico studies to cope with the complexity of cardiac models. However, all these studies are based on current-shape parameters via curve fitting, or characteristic curves/statistics from additional curve fitting applied to estimated currents. In addition, only population-average models are derived and calibrated via learning the entire dataset for each healthy and diseased group. Little work has been done to leverage computer models to calibrate and delineate kinetic dynamics of K\textsubscript{v} isoforms with multiple voltage-clamp responses simultaneously.

\subsection{Congenital Disorders of Glycosylation}
Protein glycosylation is one of the most abundant and diverse forms of co/posttranslational modifications that impact essential protein functions, such as modulation of receptor or ion channel activities \citep{ohtsubo2006glycosylation,ednie2012modulation}. A growing number of studies have shown the association between altered glycosylation and heart diseases, such as dilated cardiomyopathy (DCM) and hypertrophic cardiomyopathy \citep{ednie2019reduced2,ohtsubo2006glycosylation}. It is reported that up to 20\% of patients with congenital disorders of glycosylation (CDG), who commonly show modest reductions in protein glycosylation, present with cardiac deficits, including idiopathic DCM \citep{marques2017cardiac}. However, uncovering the underlying pathological mechanisms still remains elusive. We have investigated how regulated glycosylation contributes to heart failure in the context of electrophysiology. Electrical signaling is orchestrated activities of a variety of ion channels and transporters. VGICs are heavily glycosylated, with ~30\% of the channel mass consisting of N-/O-linked glycans \citep{ednie2012modulation}. Glycosylation is a multi-step process and usually ends with sialic acid added. We reported that a saturating, electrostatic effect of negatively charged sialic attached to the terminal of N-/O-glycan branches significantly altered electrical signaling in Na\textsubscript{v} \citep{ednie2013sialicNav1,ednie2015sialicNav2} as well as K\textsubscript{v} \citep{ednie2015sialicKv}. Computational modeling has been used to further investigate the functional role of reduced sialylation in Na\textsubscript{v} and K\textsubscript{v} \citep{du2015statistical,du2017silico}.

\section{Research Methodology}
\subsection{Computer Models of Potassium Channel Isoforms}
Because of the relative ease at which their genetics can be altered, mice have been useful models for studying cardiac electric signaling \citep{milani2014small,nerbonne2004studying}. In ventricular and atrial mouse cardiomyocytes, there are three major components of I\textsubscript{Ksum}: rapidly inactivating transient outward K\textsuperscript{+} currents (I\textsubscript{Kto,f} and/or I\textsubscript{Kto,s}), which are conducted through K\textsubscript{v}4.2 and K\textsubscript{v}1.4 respectively, slowly inactivating delayed rectifier K\textsuperscript{+} currents (I\textsubscript{Kslow1} and I\textsubscript{Kslow2}), which are conducted through K\textsubscript{v}1.5/K\textsubscript{v}2.1, and the non-inactivating steady-state K\textsuperscript{+} current (I\textsubscript{Kss}) \citep{xu1999four}. I\textsubscript{Kto,s} (K\textsubscript{v}1.4) is primarily found in septal cardiomyocytes \citep{bondarenko2004computer}. In this study, we only include I\textsubscript{Kto,f} in I\textsubscript{Kto}, because our in-vitro studies focused on the ventricular apex myocytes. However, I\textsubscript{Kto} can be easily modified according to the region of ventricular cardiomyocytes. We model these four dominant K\textsuperscript{+} currents: I\textsubscript{Kto}, I\textsubscript{Kslow1}, I\textsubscript{Kslow2}, and I\textsubscript{Kss}. Figure~\ref{fig:kcurrent_example}(b) illustrates the shape of the primary K\textsuperscript{+} currents and their contribution to the whole-cell I\textsubscript{K} trace (I\textsubscript{Ksum}) given the protocol in Figure~\ref{fig:kcurrent_example}(a) that applies 0 mV voltage pulse from holding potential -70 mV from time $t^{(h)}$ to $t^{(e)}$. Figure~\ref{fig:kcurrent_example}(c) and (d) show a range of voltage steps from -30 mV to 50 mV in 10 mV increments and consequent changes in I\textsubscript{Ksum}. Because of their shape, I\textsubscript{Kss} is assumed as a constant and the other currents an exponential function in the traditional data-driven curve-fitting approach. In this approach, if all four current models are included, I\textsubscript{Ksum} is defined by \eqref{eq:expl_fitting} (i.e., tri-exponential fitting) with the shape parameters such as amplitude $\mathrm{A}_{i}$ and time constant $\tau_{i}$ for $i \in \{\mathrm{Kto}, \mathrm{Kslow1}, \mathrm{Kslow2}, \mathrm{Kss}\}$. In some cases, it is reduced to a bi-exponential function combining I\textsubscript{Kslow1} and I\textsubscript{Kslow2}. 
\begin{figure}[!ht]
    \centering
    \includegraphics{figs/kcurrent_example.pdf}
    \caption{Example of voltage-clamp protocol and K\textsuperscript{+} current traces. (a) Clamp-voltage pulse of 0 mV from the holding potential -70 mV. (b) Dominant K\textsuperscript{+} currents and their contributions to I\textsubscript{Ksum}. (c) Series of clamp-voltage pulses (-30-50 mV in 10 mV increments) from the holding potential -70 mV and (d) consequent I\textsubscript{Ksum} traces.}
    \label{fig:kcurrent_example}
\end{figure}
\begin{equation}
    \label{eq:expl_fitting}
    \mathrm{I}_{\mathrm{Ksum}} = \mathrm{A}_{\mathrm{Kto}}e^{-t/\tau_{\mathrm{Kto}}} + \mathrm{A}_{\mathrm{Kslow1}}e^{-t/\tau_{\mathrm{Kslow1}}} + \mathrm{A}_{\mathrm{Kslow2}}e^{-t/\tau_{\mathrm{Kslow2}}} +  \mathrm{A}_{\mathrm{Kss}}.   
\end{equation}

We developed mouse K\textsubscript{v} models based on \citep{bondarenko2014compartmentalized,asfaw2020compartmentalized}, using the Hodgkin-Huxley modeling scheme that has been used for various species, such as humans \citep{ten2004model} and rabbits \citep{mahajan2008rabbit}. This type of model consists of two gating variables controlling the channel conductance, and its canonical form is defined by \eqref{eq:hh_ex}
\begin{equation}
    \label{eq:hh_ex}
    \mathrm{I}_{\mathrm{K}} = G_{\mathrm{K}}a^{n}i^{m}(V-E_{\mathrm{K}}),
\end{equation}
where $G_{\mathrm{K}}$ is the maximum conductance, $a^{n}$ and $i^{m}$ are the gating variables for $n,m \in \mathbb{N}$, $V$ is the transmembrane potential, and $E_{\mathrm{K}}$ is the K\textsuperscript{+} Nernst potential. $V-E_{\mathrm{K}}$ implies the driving force of the ion movement. Important components in \eqref{eq:hh_ex} are the gating variables $a$ and $i$, representing the fraction of activation and recovery from inactivation of the channel where $a,i \in [0,1]$. These processes are governed by first-order kinetics and voltage-dependent transition rates $\alpha$ and $\beta$. $\alpha$ is the rate at which a gate in a closed state opens, whereas $\beta$ is the rate at which a gate in an open state closes. Equation~\eqref{eq:gv} shows a schematic relationship of this gating kinetics. $n$ and $m$ represent the numbers of activation and inactivation gates, which are dependent on a particular K\textsubscript{v} isoform and determined by specific kinetics and properties of that channel. In general, the number of gates depends on the complexity of the channel behavior and the level of detail required in the model. 
\begin{align}
    \label{eq:gv}
    (1-a)&\xrightleftharpoons[\beta_{a}]{\alpha_{a}}a & (1-i)&\xrightleftharpoons[\beta_{i}]{\alpha_{i}}i.
\end{align}

These two biophysical processes $a$ and $i$ can be expressed using differential equations in two ways \eqref{eq:1stkinetic1} and \eqref{eq:1stkenetic2}
\begin{align}
    \label{eq:1stkinetic1}
    \frac{da}{dt} &=\alpha_{a}(1-a)-\beta_{a}a   &\frac{di}{dt} &=\alpha_{i}(1-i)-\beta_{i}i \\
    \label{eq:1stkenetic2}
    \frac{da}{dt} &= \frac{a_{\infty}-a}{\tau_{a}}  &\frac{di}{dt} &= \frac{i_{\infty}-i}{\tau_{i}},
\end{align}
where $a_{\infty}$ and $i_{\infty}$ are the steady-state values to which $a$ and $b$ converge; $\tau_{a}$ and $\tau_{i}$ are time constants determine the convergence speed defined by
\begin{align}
    a_{\infty} &= \frac{\alpha_{a}}{\alpha_{a}+\beta_{a}} & i_{\infty} &=  \frac{\alpha_{i}}{\alpha_{i}+\beta_{i}} \\
    \tau_{a} &= \frac{1}{\alpha_{a}+\beta_{a}} & \tau_{i} &= \frac{1}{\alpha_{i}+\beta_{i}}.
\end{align}
The steady-state values and time constants can be defined directly by functions of voltage $V$ without transition rates in some cases. These voltage-dependent functions, such as transition rates, steady states, or time constants, have parameters $p_{s}$ that control the behavior of the kinetics of an ion channel.
